Analysis of NLA17 Dissolved Gas Indicator

### 2023-06-21 scripts/dgIndicatorAnalysis.Rmd
Incremental progress on manuscript.  Added discussion of N2O saturation vs lake size
manuscript progress

all_data <- all_predictions %>%
  mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
           (k600.cm.h * (24/100)), #h->day. cm->m
         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
           (1000*1000*1000)) #mg-g, g-kg, kg-Mg

@@

The total N2O flux from US lakes and reservoirs is minor compared to other anthropogenic N2O sources
```{r}
all_predictions %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x=e.n2o.mg.m2.d)) + 
  geom_density() +
  geom_rug() + 
  facet_wrap(~WSA9, scales = "free")

knitr::include_graphics((paste0(Sys.getenv("USERPROFILE"), 
                                    "/Environmental Protection Agency (EPA)/",
                                    "ORD NLA17 Dissolved Gas - Documents/",
                                    "EDA/en2oLiterature.tiff")))

## Comparison to US GHG inventory
# top sources from inventory report in kt / year
n2o <- data.frame(kt = c(1156,88,84,66,60,34,18,14,7),
                     source = c("Agricultural Soil Management",
                                "Wastewater Treatment",
                                "Stationary Combustion",
                                "Manure Management",
                                "Mobile Combustion",
                                "Nitric Acid Production",
                                "Adipid Acid Production",
                                "N2O from Product Uses",
                                "Composting"))

@@

# flux by ecoregion
# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f), 3), 
    LCL = round(quantile(mean_f, probs = 0.025), 3),
    UCL = round(quantile(mean_f, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oSatRatioByEco.tiff")

# point and line range by size and ecoregion
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f.n2o.Mg.y = mean(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("mean N2O flux") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion, scales = "free")
ggsave("output/figures/n2oFluxBySizeEco.tiff", height = 5.4, width = 9.6)

# bar plot by size category and ecoregion  
all_predictions %>% 
  filter(.draw == 1) %>% 
  group_by(WSA9, size_cat2) %>% 
  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
  ggplot(aes(x=size_cat2, y=total.n2o)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~WSA9, scales = "free")

@@

>>>>

all_predictions <- all_predictions %>%
  mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
           (k600.cm.h * (24/100)), #h->day. cm->m
         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
           (1000*1000*1000)) #mg-g, g-kg, kg-Mg

@@

The total N2O flux from US lakes and reservoirs is minor compared to other anthropogenic N2O sources
```{r}
# Compare to literature
# Read data from Webb et al. 2019
webb <- readr::read_csv(paste0(Sys.getenv("USERPROFILE"), 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "literature/n2o/Webb et al. 2019 PNAS/",
                               "SI_data.csv")) %>%
  mutate(e.n2o.mg.m2.d = N2O.umol.m2.d * (44/1000), # umol->mg
         source = "webb")

# Read data from Deemer et al. 2016
deemer <- readxl::read_xls(paste0(Sys.getenv("USERPROFILE"), 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "literature/n2o/Deemer et al. 2016 Bioscience SI.xls"),
                           sheet = "Supplementary_Reservoir_GHG_Dat",
                           skip = 1) %>%
  mutate(e.n2o.mg.m2.d = `mg N2O-N m-2 d-1` * (44/28),
         source = "deemer")

# combine NLA and lit data
all_dat <- rbind(all_predictions %>% 
  filter(.draw == 1) %>% # just one iteration
  mutate(source = "nla") %>%
  select(source, e.n2o.mg.m2.d),
  webb %>% select(source, e.n2o.mg.m2.d),
  deemer%>% select(source, e.n2o.mg.m2.d))

# dot plot of N2O emission rates by source
ggplot(all_dat, aes(source, e.n2o.mg.m2.d)) +
  geom_point(position = "jitter")

# summary statistics by source
all_dat %>%
  split(f = list(.$source)) %>% # split into named lists
  map(~ summary(.x)) # apply summary function to each list element

## Comparison to US GHG inventory
# top sources from inventory report in kt / year
n2o <- data.frame(kt = c(1156,88,84,66,60,34,18,14,7),
                     source = c("Agricultural Soil Management", # this includes leaching and runoff
                                "Wastewater Treatment",
                                "Stationary Combustion",
                                "Manure Management",
                                "Mobile Combustion",
                                "Nitric Acid Production",
                                "Adipid Acid Production",
                                "N2O from Product Uses",
                                "Composting"))

@@

# flux by ecoregion
# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f), 3), 
    LCL = round(quantile(mean_f, probs = 0.025), 3),
    UCL = round(quantile(mean_f, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("N2O flux (Mg N2O)") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oFuxByEco.tiff")

# point and line range by size and ecoregion: facetted
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("mean N2O flux (Mg N2O year-1)") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion, scales = "free")
ggsave("output/figures/n2oFluxBySizeEco.tiff", height = 5.4, width = 9.6)

# point and line range by size and ecoregion: single plot
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
  mutate(unique = paste(WSA9, size_cat, "_")) %>%
  # mutate(ecoregion = fct_reorder(ecoregion, )) %>%
  ggplot( aes( y = size_cat, x = estimate ) ) +
  geom_point() +
  geom_linerange( aes( xmin = LCL, xmax = UCL)) +
  geom_vline(xintercept = 0, color='blue') +
  xlab(expression(N[2]*O~flux~(Mg~year^{-1}))) +
  #coord_flip() + 
  theme_bw() +
  facet_grid(WSA9~., scales = "free")
ggsave("output/figures/n2oFluxBySizeEcoVerticalPanel.tiff", height = 5.7, width = 2.5)

# bar plot by size category and ecoregion  
all_predictions %>% 
  filter(.draw == 1) %>% 
  group_by(WSA9, size_cat) %>% 
  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
  ggplot(aes(x=size_cat, y=total.n2o)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~WSA9, scales = "free")

@@

# N2O flux vs total surface area in each ecoregion
all_predictions %>%
  group_by(WSA9, WSA9_NAME, .draw) %>%
  summarise(fn2o_Mg = sum(f.n2o.Mg.y),
             total_sa = sum(area_ha)) %>%
  summarise( post_med_fn2o = round(median(fn2o_Mg), 3), 
             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
             total_sa = mean(total_sa)) %>%
  ggplot(aes(total_sa, post_med_fn2o)) +
  geom_point()
